

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library("readxl")
library("dplyr")
library("tidyverse")
library(psych)

getwd()

data <- read_xlsx("/Users/AB/Desktop/Cyprus-Research-Program/Clean DataLogs for impactors and Alphasense(cor_Dates+ Match Alphasense)_Updated AM 30-10-22  (1).xlsx")


View(data)

# Trimming 
df <- data %>% select(`PM (PM2.5,PM10)`, Average, Median, 
                     `PM MASS (μg/m3)`, `Impactor Location (Indoor/Outdoor)`,
                     Flag, `INTERVENTION GROUP`, `Dust(D)/ NonDust(ND)`)

#provisionally ignore subjects with missing values 
df <- df[which(!is.na(df$Median)),] 

df <- df[which(!is.na(df$Average)),] 

#making PM MASS values numeric 
df$`PM MASS (μg/m3)` <- as.numeric(df$`PM MASS (μg/m3)`)
df <- df[which(!is.na(df$`PM MASS (μg/m3)`)),] #removing N/As 

# Create PM2.5 data set. Removing flag 2  
PM2.5 <- df %>% filter(`PM (PM2.5,PM10)` == "PM2.5", Flag != "2", 
                       Average != "35.26")

# creating PM2.5 df without outliers 
boxplot(PM2.5$Average)
boxplot(PM2.5$`PM MASS (μg/m3)`)

PM2.5_no_outlier <- PM2.5 %>% filter(Average != "35.2582041", 
                                     `PM MASS (μg/m3)`!= 47.255555) 
boxplot(PM2.5_no_outlier$Average)
boxplot(PM2.5_no_outlier$`PM MASS (μg/m3)`)

rename(PM2.5_no_outlier, AveragePM2.5 = Average) #remaining column 

# Create PM10 data set. Removing flag 2  
PM10 <- df %>% filter(`PM (PM2.5,PM10)` == "PM10", Flag != "2")

# creating PM10 df without outliers 
boxplot(PM10$Average) 
boxplot(PM10$`PM MASS (μg/m3)`)

PM10_no_outlier <- PM10 %>% filter(Average < 50, `PM MASS (μg/m3)`!= "175.291046")

boxplot(PM10_no_outlier$Average)
boxplot(PM10_no_outlier$`PM MASS (μg/m3)`)
```


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# data visualization
library("ggplot2")
library("ggthemes")
library("ggrepel")

hist1 <- PM2.5_no_outlier %>% 
  ggplot(aes(`PM MASS (μg/m3)`, ..density..)) + 
  geom_histogram(binwidth = 1, color="black", fill="navy") +
  xlab("PM MASS (μg/m3) HI") 

hist2 <- PM2.5_no_outlier %>% 
  ggplot(aes(log(`PM MASS (μg/m3)`), ..density..)) + 
  geom_histogram(binwidth = 0.25, color="black", fill="navy") + 
  xlab("Log of PM MASS (μg/m3) HI") 

hist3 <- PM2.5_no_outlier %>% 
  ggplot(aes(Average, ..density..)) + 
  geom_histogram(binwidth = 0.5, color="black", fill="navy") + 
  xlim(0, 20)
  xlab("Average PM2.5(μg/m3) Alphasensor") 

hist4 <- PM2.5_no_outlier %>% 
  ggplot(aes(log(Average), ..density..)) + 
  geom_histogram(binwidth = 0.25, color="black", fill="navy") + 
  xlab("Log Average PM2.5(μg/m3) Alphasensor")

hist5 <- PM10_no_outlier %>% 
  ggplot(aes(Average, ..density..)) + 
  geom_histogram(binwidth = 5, color="black", fill="navy") + 
  xlab("Average PM10 (μg/m3) Alphasensor") 

hist6 <- PM10_no_outlier %>% 
  ggplot(aes(log(Average), ..density..)) + 
  geom_histogram(binwidth = 0.25, color="black", fill="navy") + 
  xlab("Log Average PM10 (μg/m3) Alphasensor") 

library("gridExtra")
grid.arrange(hist1, hist2, hist3, hist4, hist5, hist6, nrow = 3)

```




### Linear Regression Models 

<br>
<br>

#### Average PM2.5 from Alphasenors predicting PM Mass from Harvard Impactor.
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(broom)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library("gridExtra")
library(sjPlot)

model1 <- lm(`PM MASS (μg/m3)` ~ Average, data=PM2.5_no_outlier)  
#PM2.5 # R^2 (coefficient of determination) is the proportion of the variance in the  
# average that is predicted by PM Mass. In other words, how well does avg. 
# predict PM Mass? 
model1
model2 <- lm(`PM MASS (μg/m3)` ~ log(Average), data=PM2.5_no_outlier)  

tab_model(model1, model2)

p1 <- PM2.5_no_outlier %>% 
    ggplot(aes(x = Average, y = `PM MASS (μg/m3)`)) + 
    geom_point(color="navy", size = 3, alpha=0.5) +  
    geom_smooth(method=lm, se=TRUE, col="black", size=1) +
    stat_regline_equation(label.x=1, label.y=31) +
    stat_cor(aes(label=..rr.label..), label.x=1, label.y=29) +
    ylab("PM Mass (μg/m3)") + 
    xlab("Alphasense PM2.5 (μg/m3)") +
    theme_classic() + 
    theme(legend.position = "none") 
p1

p2 <-  PM2.5_no_outlier %>% 
    ggplot(aes(x = log(Average), y = `PM MASS (μg/m3)`)) + 
    geom_point(color="navy", size = 3, alpha=0.5) +  
    geom_smooth(method=lm, se=TRUE, col="black", size=1) +
    stat_regline_equation(label.x=0.2, label.y=20) +
    stat_cor(aes(label=..rr.label..), label.x=0.2, label.y=17) +
    ylab("Impactor PM Mass (μg/m3)") + 
    xlab("Alphasense Log PM2.5 (μg/m3)") +
    theme_classic() + 
    theme(legend.position = "none") 

p2

grid.arrange(p1, p2, nrow = 2)


```
<br>
<br>
<br>

#### Y = Fitted Alphasense PM2.5 values, X = Actual Alphasense PM2.5 Measurement 

```{r echo=FALSE}
tab1 <- predict(model2, interval = "confidence")

PM2.5_fit <- bind_cols(tab1, PM2.5_no_outlier)

fitmod1 <- lm(fit ~ Average, data = PM2.5_fit)

tab_model(fitmod1)
```

<br>
<br>
<br>

#### Average PM2.5 from Alphasenors predicting PM Mass from Harvard Impactor adjusting for dust and intervention status. 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# adding covariates
model3 <- glm(`PM MASS (μg/m3)`~ log(Average) + `INTERVENTION GROUP` + 
               `Dust(D)/ NonDust(ND)`, data=PM2.5_no_outlier)

# adding interaction term
model4 <- glm(`PM MASS (μg/m3)`~ log(Average) + `INTERVENTION GROUP` +
                `Dust(D)/ NonDust(ND)` + `INTERVENTION GROUP` * `Dust(D)/ NonDust(ND)`,
              data = PM2.5_no_outlier)

tab_model(model3, model4)
``` 


```{r eval=FALSE, include=FALSE}
library(car) 

#plotting the regression line of adjusted models for PM2.5 
avPlots(model3, terms=~., intercept = FALSE, layout = NULL, grid = FALSE, 
        ellipse=FALSE, pt.wts = FALSE, lwd =2)

avPlots(model4, terms=~., intercept = FALSE, layout = NULL, grid = FALSE, 
        ellipse=FALSE, pt.wts = FALSE, lwd =2)
```

<br>
<br>
<br>

#### Average PM10 from Alphasenors predicting PM Mass from Harvard Impactor.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

model5 <- lm(`PM MASS (μg/m3)` ~ Average, data=PM10_no_outlier)  

model6 <- lm(`PM MASS (μg/m3)` ~ log(Average), data=PM10_no_outlier)  

tab_model(model5, model6)

plot3 <- PM10_no_outlier %>% 
  ggplot(aes(x = Average, y = `PM MASS (μg/m3)`)) + 
  geom_point(color="navy", size = 3, alpha=0.5) +
  geom_smooth(method=lm, se=TRUE, col="black", size=0.5) +
  stat_regline_equation(label.x=2, label.y=53) +
  stat_cor(aes(label=..rr.label..), label.x=2, label.y=49) +
  ylab("PM Mass (μg/m3)") + 
  xlab("Alphasense PM10 (μg/m3)") +
  theme_classic() + 
  theme(legend.position = "none") 

plot3

plot4 <- PM10_no_outlier %>% 
  ggplot(aes(x = log(Average), y = `PM MASS (μg/m3)`)) + 
  geom_point(color="navy", size = 3, alpha=0.5) +
  geom_smooth(method=lm, se=TRUE, col="black", size=0.5) +
  stat_regline_equation(label.x=0.5, label.y=35) +
  stat_cor(aes(label=..rr.label..), label.x=0.5, label.y=29) +
  ylab("Impactor PM Mass (μg/m3)") + 
  xlab("Alphasense Log PM10 (μg/m3)") +
  theme_classic() + 
  theme(legend.position = "none")

plot4

grid.arrange(plot3, plot4, nrow = 2)

together <- grid.arrange(p2, plot4, nrow = 1)
together
# Save the plot as an image (PNG format)
ggsave("colocation_prescaling.png", together, width = 12, height = 5, units = "in", dpi = 300)

colo_prescale_nolog <- grid.arrange(p1, plot3, nrow = 1)
colo_prescale_nolog

ggsave("colo_prescle_nolog.png", colo_prescale_nolog, width = 12, height = 5, units = "in", dpi = 300)
```

### Agreement post-scaling 
```{r}

data2 <- read_csv("/Users/AB/Desktop/Cyprus-Research-Program/Azariah_data.csv")

# filter intervention variable = no intervention and full intervention
data3 <- data2 %>% 
  filter(INTERV == 0 | INTERV == 2) %>% 
  filter(FLAG == 0) # remove flags an dust days
unique(data3$INTERV)

# create 0, 1 filter variable 
data3$filt <- ifelse(data3$INTERV==0, 0, 1)
unique(data3$filt)

# scaling
data3$ahour <- data3$Hour + 1 

data3$spm25 <- data3$aPM25 * data3$scale_ratio25      
data3$lspm25 <- log(data3$spm25)

data3$spm10 <- data3$aPM10 * data3$scale_ratio10      
data3$lspm10 <- log(data3$spm10)

data3$lipm10 <- log(data3$iPM10)
data3$lipm25 <- log(data3$iPM25)

data3$lopm10 = log(data3$oPM10)
data3$lopm25 = log(data3$oPM25)

data3$r25 <- (data3$spm25 / data3$oPM25)
data3$r10 <- (data3$spm10 / data3$oPM10)

data3$lr25 <- log(data3$r25)
data3$lr10 <- log(data3$r10)

View(data3)


#filtering data for graphs 
data4 <- data3 %>% filter(spm10<60, spm25<40)

logscaled1 <- data4 %>% 
  ggplot(aes(x = lspm25, y = lr25)) + 
  geom_point(color="navy", size = 3, alpha=0.5) +
  geom_smooth(method=lm, se=TRUE, col="black", size=0.5) +
  stat_regline_equation(label.x=-1.25, label.y=3) +
  stat_cor(aes(label=..rr.label..), label.x=-1.25, label.y=2.5) +
  ylab("Log Adjusted PM Mass (μg/m3)") + 
  xlab("Log Adjusted Alphasense PM25 (μg/m3)") +
  theme_classic() + 
  theme(legend.position = "none")
logscaled1

install.packages("ggtext")
library(ggtext)


logscaled2 <- data4 %>% 
  ggplot(aes(x = lspm10, y = lr10)) + 
  geom_point(color="navy", size = 3, alpha=0.5) +
  geom_smooth(method=lm, se=TRUE, col="black", size=0.5) +
  stat_regline_equation(label.x=-1.5, label.y=4) +
  stat_cor(aes(label=..rr.label..), label.x=-1.5, label.y=3.4) +
  ylab("Log Adjusted PM Mass (μg/m^3)") + 
  xlab("Log Adjusted Alphasense PM[10] (μg/m^3)") +
  theme_classic() + 
  theme(legend.position = "none")
logscaled2


forpub <- grid.arrange(
  p1, 
  plot3, 
  logscaled1, 
  logscaled2,
  nrow = 2,
  bottom = textGrob(
    "PM Mass refers to Harvard Impactor concentrations",
  gp = gpar(fontface = 3, fontsize = 9),
  hjust = 1,
  x = 1
  )
)


ggsave("forpub.png",forpub, width = 12, height = 6, units = "in", dpi = 300)

```


#### Y = Fitted Alphasense PM10 values, X = Actual Alphasense PM10 Measurement 
```{r echo=FALSE}
tab2 <- predict(model6, interval = "confidence")

PM10_fit <- bind_cols(tab2, PM10_no_outlier)

fitmod2 <- lm(fit ~ Average, data = PM10_fit)

tab_model(fitmod2)
```
<br>
<br>
<br>

##### Average PM10 from Alphasenors and PM Mass from Harvard Impactor. Adjusting for dust and intervention status. 
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#adding covariates 
model6 <- glm(`PM MASS (μg/m3)`~ log(Average) + `INTERVENTION GROUP` + 
               `Dust(D)/ NonDust(ND)`, data=PM10_no_outlier)

# adding interaction term
model7 <- glm(`PM MASS (μg/m3)`~ log(Average) + `INTERVENTION GROUP` +
                `Dust(D)/ NonDust(ND)` + `INTERVENTION GROUP` * `Dust(D)/ NonDust(ND)`,
              data = PM10_no_outlier)

tab_model(model6, model7)
```

```{r eval=FALSE, include=FALSE}
#plotting the regression line of adjusted models for PM10
avPlots(model4, terms=~., intercept = FALSE, layout = NULL, grid = FALSE, 
        ellipse=FALSE, pt.wts = FALSE, lwd =2)
```


```{r eval=FALSE, include=FALSE}
#### Plotting Observed vs. Predicted Values 
ggplot(PM2.5_no_outlier, aes(x=predict(model2), 
           y= (PM2.5_no_outlier$`PM MASS (μg/m3)`))) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  labs(x='Predicted Log Average PM2.5 Values', y='Actual Log Average PM2.5 Values', title='Predicted vs. Actual Values')

```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#### Particulate Matter Measurements over Time 
library(dplyr)
library(readr)
library(ggplot2)
library(ggthemes)

df2 <- data %>% select(`PM (PM2.5,PM10)`, Average, Median, 
                     `PM MASS (μg/m3)`, `Impactor Location (Indoor/Outdoor)`,
                     Flag, `Sampling time (min) With Correction timestamps`) %>%
  filter(Flag != "2")

df2 <- df2[which(!is.na(df2$`PM MASS (μg/m3)`)),]   # remove missing values 
df2 <- df2[which(!is.na(df2$Average)),] 
df2 <- df2[which(!is.na(df2$Median)),] 

write.csv(df2, "~/Desktop/df2.csv")
df2 <- read_csv("~/Desktop/df2_edit.csv")

#making PM MASS values numeric 
df2$`PM MASS (μg/m3)` <- as.numeric(df2$`PM MASS (μg/m3)`)

df2 %>% ggplot(aes(x = `Sampling time (min) With Correction timestamps`, 
                   `PM MASS (μg/m3)`, y = Average, color = `PM (PM2.5,PM10)`)) + 
  labs(color = "PM") +
  stat_smooth(method = loess, se = FALSE) +
  ylim(0, 40) +
  ylab("Average Particulate Matter (μg/m3)") +
  xlab("Sampling Time (minutes)") + 
  theme_clean()
```

Calibration curve and predict hourly. 
Use the new column that Antonis made. 
After you predict the PM based on the Alphasense.. then you don't need to use the averages.
house number 
day 
time 
dust 
intervent
pred

